---
title: "StatsClass_Analysis"
author: "Sydney Horan"
date: '2023-04-15'
output: html_document
---
Installing and loading required packages

```{r error=FALSE, warning=FALSE, message=FALSE}
#Packages... I don't know if all of these are useful but just in case
library(car) #ANOVA
library(lme4) #GLMM and GLM
library(glmmTMB) #Poisson lots of zeros
library(MASS) #glm.nb negative binomial lots of zeros 
library(tidyverse)
library(sjPlot)
library(vegan)
library(cowplot)
library(ggplot2)
library(ggpubr)
library(ggbeeswarm)
library(sjstats)
library(showtext) #Adding some more fonts
font_add_google("Roboto", "Roboto")

data <- read.csv("Horan_CT2022_NestlingLevelData_SMI_Apr2023.csv") #Reading in the dataset (nestling)
nestdata <- read.csv("Horan_CT2022_NestLevelData.csv") #Reading in dataset (nest level)
```
#### Question 1 - Effect of treatment on nest temperature
* Fixed Effect: Treatment
* Response Variable:Temperature
* Random Effect: None
* Co-variate: Julian Day (ie. time of year)
```{r}
#Step 1: Test for Normality 
shapiro.test(nestdata$NestTemperature_C)
#W = 0.93054, p-value = 0.521
# The p-value is greater than 0.05. The data is normally distributed.

qqnorm(nestdata$NestTemperature_C)
qqline(nestdata$NestTemperature_C, col = "gray")

#Step 2: Creating the model
model1 <- glm(NestTemperature_C~Treatment+JulianHatchDay, family=gaussian, nestdata)
Anova(model1)
#Response: NestTemperature_C
             #  LR Chisq Df Pr(>Chisq)    
#Treatment        30.638  1  3.109e-08 ***
#JulianHatchDay   15.972  1  6.427e-05 ***
summary(model1)

#Step 2: Visualizing
p1<- ggplot(nestdata, aes(x=Treatment, y=NestTemperature_C)) + 
  geom_boxplot()+
  geom_jitter(position=position_jitter(0.2))+
  labs(title="Effect of Treatment on Temperature",x="Treatment", y = "Average Nest Temperature (°C)")+
  theme_classic()

#plot_model(model1, type="pred", terms = c("JulianHatchDay", "Treatment"))

table1 <- Anova(model1)
kable(table1, digits = 3) # the digits argument controls rounding
```

Trying to pretty up the box plot..

```{r}
tiff("tempfig.tiff", units="in", width=6, height=4, res=500)
treatmentbp <- nestdata %>% 
  ggplot(aes(x = Treatment, y = NestTemperature_C)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill=factor(Treatment)), shape=21, colour="black", size=3, stroke=1, width=.2) +
  scale_fill_manual(values = c("#227c9d", "#fe6d73")) +
  labs(x = "",
       y = "Average Nest Temperature (°C)\n") +
  theme(
    plot.title = element_text(size = 18, face = "bold", family="Roboto", hjust = 0.5),
    axis.text.x = element_text(size = 13, face = "bold", angle = 0, family="Roboto"),
    axis.title=element_text(size=16, face="bold", family="Roboto"),
    axis.text.y = element_text(size = 12),
    legend.position = "None",
    strip.text = element_text(size = 16),
    panel.background = element_blank(), #changes background color. remove "#" before this if you would prefer a white background
    panel.border=element_rect(fill=NA),
    strip.background = element_rect(fill = NA, colour = "black"),
    panel.grid.major.y = element_line(color="gray62",linetype="dotted"),
          panel.grid.minor.y = element_line(color="gray64",linetype="dotted"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

treatmentbp + ggtitle("Effect of Treatment on Nest Temperature")
```


#### Question 2 - Effect of treatment on immune gene expression
* Fixed Effect: Treatment
* Response Variable: IL-6, HSP70
* Random Effect: NestID
* Co-variate: Julian Day (ie. time of year), Sex, Age

##### Testing expression of IL-6 in response to temperature

```{r}
#Step 1: Test for Normality 
shapiro.test(data$IL.6_GAPDH)
#W = 0.64653, p-value = 2.066e-07
# The p-value is less than 0.05. The data is not normally distributed. Let's log10 transform the data.

shapiro.test(data$IL.6log_GAPDH)
#W=0.96081, p-value = 0.3065
# The p-value is more than 0.05. The data is now normally distributed.

model2 <- lmer(IL.6log_GAPDH~Treatment+Sex+Age+JulianDate+(1|Nest.ID), data=data)
Anova(model2)
# Response: IL.6log_GAPDH
#            Chisq Df Pr(>Chisq)  
# Treatment  2.9240  1    0.08727 .
# Sex        2.3222  1    0.12754  
# Age        0.1970  1    0.65714  
# JulianDate 3.2181  1    0.07283 .
summary(model2)

#Running without covariates, out of curiosity
#model2a <- lmer(IL.6log_GAPDH~Treatment+(1|Nest.ID), data=data)
#Anova(model2a)
  # Chisq Df Pr(>Chisq)
#Treatment 2.1226  1     0.1451
#Looks like they are definitely necessary.

```

```{r}
# Messing around with visualizing the data. Here is a plot of effect sizes.
sjPlot::plot_model(model2)
sjPlot::plot_model(model2, 
                   axis.labels=c("Julian Date", "Age", "Sex (M)", "Treatment (Control)"),
                   show.values=TRUE, show.p=TRUE,
                   title="Effect of Heat on IL-6 Expression")

#Making the above plot into a table 
sjPlot:: tab_model(model2)
sjPlot::tab_model(model2, 
                  show.re.var= TRUE, 
                  pred.labels =c("(Intercept)", "Treatment (Control)", "Sex (M)", "Age", "Julian Date"),
                  dv.labels= "Effect of Heat on IL-6 Expression")

#Getting the effects of the treatment fixed effect specifically
effects_heat <- effects::effect(term= "Treatment", mod = model2)
summary(effects_heat)
#Treatment effect
#Treatment
#       Heat     No Heat 
#-0.37609486  0.02782042 

# Lower 95 Percent Confidence Limits
#Treatment
#      Heat    No Heat 
#-0.7130327 -0.3055634 

# Upper 95 Percent Confidence Limits
#Treatment
#       Heat     No Heat 
# -0.03915705  0.36120420 

#Putting this into a dataframe to graph potentially
x_heat <- as.data.frame(effects_heat)
```


###### Graphing treatment vs IL-6 expression mixed effects model

```{r}
# data = raw data file
#x_heat = effects of treatment statistical values from above model
plot1 <- ggplot(data, aes(x = Treatment, y = IL.6log_GAPDH, color = Treatment)) +
  geom_beeswarm(cex = 4, size=2) +
  scale_color_brewer(palette = "Pastel1") +
  scale_y_continuous(limits=c(-1,1), breaks = seq(-1, 1, by = 1)) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", width=0.05, alpha=0.7, color="black") +
  stat_summary(fun=mean, geom="point", color="black") +
  labs(x = "",
       y = "Relative Gene Expression") +
  theme(
    plot.title = element_text(size = 16, face = "bold", family="Roboto", hjust = 0.5),
    axis.text.x=element_text(size = 16, angle = 0, family="Roboto", face="bold"),
    axis.title=element_text(size=16, family="Roboto", face="bold"),
    axis.text.y = element_text(size = 16, family="Roboto", face="bold"),
    legend.position = "None",
    strip.text = element_text(size = 16),
    panel.background = element_blank(),
    panel.border=element_rect(fill=NA),
    strip.background = element_rect(fill = NA, colour = "black"),
  )
plot1 #<- #plot1 + ggtitle("Effect of Heat on IL-6 Expression")
#plot1 + geom_point(data=x_heat, aes(x=Treatment, y=fit), color="black") # If you run this, it will add a data point that's a mean as predicted by the model plot. I cannot get it to add error bars, just this point.


#QUESTION: How do I get error bars to work on this when I have the SE and confidence intervals already in a dataframe (x_heat)? Or am I going about this wrong?
```


#### Testing New Code for graphing Model Output - 24 Apr 2023

```{r}

# simple way to plot and check the model
plot_model(model2, type = "pred", terms = c("Treatment")) 
#IT WORKED!!

#putting into a dataframe just in case
model2df <- get_model_data(model2, type = "pred", terms = c("Treatment"))
model2df <- as.data.frame(model2df)

write.csv(model2df, "~/Documents/StatsClass_2023/StatsClass/FinalProject/model2_data.csv", row.names=FALSE)

# Predicted model values shown in the figure:
```


##### Testing expression of HSP70 in response to temperature

```{r}
#Step 1: Test for Normality 
shapiro.test(data$HSP70_GAPDH)
#W = 0.76891, p-value = 1.444e-05
# The p-value is less than 0.05. The data is not normally distributed. Let's log10 transform the data.

shapiro.test(data$HSP70log._GAPDH)
#W=0.96081, p-value = 0.2117
# The p-value is more than 0.05. The data is now normally distributed.

model3 <- lmer(HSP70log._GAPDH~Treatment+Sex+Age+JulianDate+(1|Nest.ID), data=data)
Anova(model3)
        #Response: HSP70log._GAPDH
         #   Chisq Df Pr(>Chisq)  
#Treatment  2.8662  1    0.09046 .
#Sex        4.9965  1    0.02540 *
#Age        2.9015  1    0.08850 .
#JulianDate 4.2318  1    0.03967 *
summary(model3)

#Out of curiosity, let's run it without the co-variates.
#model3a <- lmer(HSP70log._GAPDH~Treatment+(1|Nest.ID), data=data)
#Anova(model3a)
  # Chisq Df Pr(>Chisq)
#Treatment 0.8148  1     0.3667
#Much sadness.

plot_model(model3, type="pred", terms = c("Sex"))
```

Let us now attempt to plot some things.

```{r}
plot2 <- ggplot(data, aes(x = Treatment, y = HSP70log._GAPDH, color = Treatment)) +
  geom_beeswarm(cex = 4, size=2) +
  scale_color_brewer(palette = "Pastel1") +
  scale_y_continuous(limits=c(-1,1), breaks = seq(-1, 1, by = 1)) +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", width=0.05, alpha=0.7, color="black") +
  stat_summary(fun=mean, geom="point", color="black") +
  labs(x = "",
       y = "Relative Gene Expression") +
  #facet_grid(. ~ UrbanOrRural, scales = "free") +
  theme(
    #axis.text.y.left = element_blank(), #remove y axis labels
    plot.title = element_text(size = 16, face = "bold", family="Roboto", hjust = 0.5),
    axis.text.x=element_text(size = 16, angle = 0, family="Roboto", face="bold"),
    axis.title=element_text(size=16, family="Roboto", face="bold"),
    axis.text.y = element_text(size = 16, family="Roboto", face="bold"),
    legend.position = "None",
    strip.text = element_text(size = 16),
    panel.background = element_blank(),
    panel.border=element_rect(fill=NA),
    strip.background = element_rect(fill = NA, colour = "black"),
  )
plot2
```

Making this plot easier to look at when I combine it with the other one-

```{r, include=FALSE}
#I am now going to remove the Y axis from this just to make it pretty when I combine the two graphs.

plot2a <- ggplot(data, aes(x = Treatment, y = HSP70log._GAPDH, color = Treatment)) +
  geom_beeswarm(cex = 4, size=2) +
  scale_color_brewer(palette = "Pastel1") +
  scale_y_continuous(limits=c(-1,1), breaks = seq(-1, 1, by = 1)) +
stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", width=0.05, alpha=0.7, color="black") +
stat_summary(fun=mean, geom="point", color="black") +
  labs(x = "",
       y = "") +
  #facet_grid(. ~ UrbanOrRural, scales = "free") +
  theme(
    #axis.text.y.left = element_blank(), #remove y axis labels
    plot.title = element_text(size = 16, face = "bold", family="Roboto", hjust = 0.5),
    axis.text.x=element_text(size = 16, angle = 0, family="Roboto", face="bold"),
    axis.title=element_text(size=16, family="Roboto", face="bold"),
    axis.text.y = element_text(size = 16, family="Roboto", face="bold"),
    legend.position = "None",
    strip.text = element_text(size = 16),
    panel.background = element_blank(),
    panel.border=element_rect(fill=NA),
    strip.background = element_rect(fill = NA, colour = "black"),
  )
plot2a #<- plot2a + ggtitle("Effect of Heat on HSP70 Expression")
```

#### Plotting effect size

```{r}
# Messing around with visualizing the data. Here is a plot of effect sizes.
sjPlot::plot_model(model3)
sjPlot::plot_model(model3, 
                   axis.labels=c("Julian Date", "Age", "Sex (M)", "Treatment (Control)"),
                   show.values=TRUE, show.p=TRUE,
                   title="Effect of Heat on HSP70 Expression")

#Making the above plot into a table 
sjPlot:: tab_model(model3)
sjPlot::tab_model(model3, 
                  show.re.var= TRUE, 
                  pred.labels =c("(Intercept)", "Treatment (Control)", "Sex (M)", "Age", "Julian Date"),
                  dv.labels= "Effect of Heat on IL-6 Expression")

#Getting the effects of the treatment fixed effect specifically
hspeffects_heat <- effects::effect(term= "Treatment", mod = model3)
summary(hspeffects_heat)
#Treatment effect
#Treatment
 #      Heat     No Heat 
#-0.29838632  0.07060448 

# Lower 95 Percent Confidence Limits
#Treatment
#      Heat    No Heat 
#-0.6082208 -0.2397373 

# Upper 95 Percent Confidence Limits
# Treatment
#      Heat    No Heat 
# 0.01144822 0.38094622 

#Putting this into a dataframe to graph potentially
x_heathsp <- as.data.frame(hspeffects_heat)
```


#### Combining into one graph

```{r, warning=FALSE}
tiff("combined.tiff", units="in", width=6, height=4, res=500)
combined <- ggarrange(plot1 + ggtitle("IL-6"), plot2a + ggtitle("HSP70"))
annotate_figure(combined,
                top = text_grob("Effect of Heat on Gene Expression", face="bold", size="20", family="Roboto")
                )
#QUESTION: The y axis scale is different between these two (though it is close!!), any way to quickly standardize it?
```


#### Question 3 - Effect of sex on immune gene expression (IL-6)
* Fixed Effect: Treatment * Sex
* Response Variable: IL-6
* Random Effect: NestID
* Co-variate: Julian Day (ie. time of year)

```{r}
model4 <- lmer(IL.6log_GAPDH~Treatment*Sex+JulianDate+(1|Nest.ID), data=data)
Anova(model4)
            #           Chisq Df Pr(>Chisq)  
#Treatment     3.0173  1    0.08238 .
#Sex           1.9854  1    0.15882  
#JulianDate    2.7796  1    0.09547 .
#Treatment:Sex 0.1150  1    0.73455  
summary(model4)

plot_model(model4, type = "pred", terms = c("Treatment", "Sex")) 
```

Plotting

```{r}
tiff("il6sexfig.tiff", units="in", width=6, height=4, res=500)
plot3 <- data %>%
  filter(!is.na(Sex)) %>%
  ggplot(aes(x = Treatment, y = IL.6log_GAPDH, color = Treatment)) +
  geom_beeswarm(cex = 4, size=2) +
  scale_color_brewer(palette = "Pastel1") +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", width=0.05, alpha=0.7, color="black") +
  stat_summary(fun=mean, geom="point", color="black") +
  facet_grid(. ~ Sex, scales = "free") +
  labs(x = "",
       y = "Relative Gene Expression (of IL-6)") +
  theme(
    plot.title = element_text(size = 16, face = "bold", family="Roboto", hjust = 0.5),
    axis.text.x = element_text(size = 13, angle = 0,  family="Roboto", face="bold"),
    axis.title=element_text(size=16, family="Roboto", face="bold"),
    axis.text.y = element_text(size = 12, family="Roboto", face="bold"),
    legend.position = "None",
    strip.text = element_text(size = 16, family="Roboto", face="bold"),
    panel.background = element_blank(),
    panel.border=element_rect(fill=NA),
    strip.background = element_rect(fill = NA, colour = "black"),
  )
  
plot3 + ggtitle("Effect of Sex on Immune Gene Expression")
#I'm not sure why it won't use the same palette as above. There's also a probably more intuitive way to visualize this.... 
```

```{r}
model5 <- lmer(HSP70log._GAPDH~Treatment*Sex+JulianDate+(1|Nest.ID), data=data)
Anova(model5)
```

```{r}
model6 <- lmer(HSP70log._GAPDH~Treatment*JulianDate+Sex+(1|Nest.ID), data=data)
Anova(model6)
```

```{r}
model7 <- lmer(IL.6log_GAPDH~Treatment+JulianDate*Sex+(1|Nest.ID), data=data)
Anova(model7)
summary(model7)
```

```{r}
tiff("hsp70sexfig.tiff", units="in", width=6, height=4, res=500)
plot4 <- data %>%
  filter(!is.na(Sex)) %>%
  ggplot(aes(x = Treatment, y = HSP70log._GAPDH, color = Treatment)) +
  geom_beeswarm(cex = 4, size=2) +
  scale_color_brewer(palette = "Pastel1") +
  stat_summary(fun.data=function(x){mean_cl_normal(x, conf.int=.683)}, geom="errorbar", width=0.05, alpha=0.7, color="black") +
  stat_summary(fun=mean, geom="point", color="black") +
  facet_grid(. ~ Sex, scales = "free") +
  labs(x = "",
       y = "Relative Gene Expression (of HSP70)") +
  theme(
    plot.title = element_text(size = 16, face = "bold", family="Roboto", hjust = 0.5),
    axis.text.x = element_text(size = 13, angle = 0,  family="Roboto", face="bold"),
    axis.title=element_text(size=16, family="Roboto", face="bold"),
    axis.text.y = element_text(size = 12, family="Roboto", face="bold"),
    legend.position = "None",
    strip.text = element_text(size = 16, family="Roboto", face="bold"),
    panel.background = element_blank(),
    panel.border=element_rect(fill=NA),
    strip.background = element_rect(fill = NA, colour = "black"),
  )
  
plot4 + ggtitle("Effect of Sex on Immune Gene Expression")
#I'm not sure why it won't use the same palette as above. There's also a probably more intuitive way to visualize this.... 
```

```{r}
juliantplot <- ggplot(data, aes(x = as.numeric(JulianDate), y = HSP70log._GAPDH)) +
  geom_point(aes(fill=Treatment), shape = 21, colour = "black", size = 3, stroke = 1, width = .2) +
  scale_fill_manual(values = c("#227c9d", "#fe6d73", "#ffcb77")) +
  #geom_smooth(color="black", method = "lm", show.legend = FALSE) +
  geom_smooth(aes(fill=factor(Treatment)), color="black", method = "lm", show.legend = FALSE) +
  #scale_color_manual(values = c("#00AFBB", "#E7B800"))+
  #scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
  xlab("\nJulian Date")+
  ylab("HSP70\n")+
  guides(fill=guide_legend(title="Aaaaaaah"))+
  theme_classic()+
  theme(
    axis.title.x = element_text(color="black", size=18),
    axis.title.y = element_text(color="black", size=18), ## , face="bold"
    strip.text = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.title = element_text(size=13)
  )
juliantplot
```

